---
title: "AE 16: Hypothesis tests for independence"
date: "October 18, 2021"
author: "STA 199"
---

```{r load-packages, warning = F, message = F}
library(tidyverse)
library(tidymodels)
```

```{r load-data}
yawn <- read_csv("yawn.csv")
```

# Learning goals

Use and understand simulation-based methods to ...

- test a claim about a population proportion
- test a claim about independence between two groups

# Part 1: Hypothesis test for a single proportion[^1]

A large university knows that about 70% of the full-time students are employed at least 5 hours per week. The members of the Statistics Department wonder if the same proportion of their students work at least 5 hours per week. They randomly sample 25 majors and find that 15 of the students work 5 or more hours each week. Use the code below to create a data frame of the results.

```{r create-df}
stats_work <- tibble(work_hours = c(rep("At least 5", 15), 
                               rep("Less than 5", 10)))
```

- State the null and alternative hypotheses. 

- Fill in the code to generate the null distribution. 

```{r prop-null}
set.seed(101821)
#null_dist <- ____ %>%
#  specify(response = ____, success = "_____") %>%
#  hypothesize(null = "point", p = ____) %>%
#  generate(reps = 100, type = "draw") %>%
#  calculate(stat = "prop")
```

- Visualize the null distribution and shade in the area used to calculate the p-value. 

```{r prop-null-viz}
# add code 
```

- Calculate the p-value. Then use the p-value to make your conclusion using a significance level of 0.05.

```{r prop-calc-p-value}
# add code
```

## Part 2: Test for independence

- First let's, watch the experiment from *Mythbusters*. 

- Let $t$ be the treatment group who saw a person yawn, $c$ be the control group who did not see anyone yawn,a nd $p$ be the proportion of people who yawned. 

## Exercise 1

We want to use simulation-based inference to assess whether or not yawning and seeing someone yawn are independent. 

- State the null and alternative hypotheses in words: 

- Select the appropriate null and alternative hypotheses written in mathematical notation: 

a. $H_0: p_t = p_c \text{ vs. }H_a: p_t < p_c$
b. $H_0: p_t = p_c \text{ vs. }H_a: p_t > p_c$
c. $H_0: p_t = p_c \text{ vs. }H_a: p_t \neq p_c$
d. $H_0: \hat{p}_t = \hat{p}_c \text{ vs. }H_a: \hat{p}_t < \hat{p}_c$
e. $H_0: \hat{p}_t = \hat{p}_c \text{ vs. }H_a: \hat{p}_t > \hat{p}_c$
f. $H_0: \hat{p}_t = \hat{p}_c \text{ vs. }H_a: \hat{p}_t \neq \hat{p}_c$

## Exercise 2

Uncomment the code to see read in the data from the class and visualize the null distribution. 

```{r sim-results}
sim_data <- read_csv("https://sta199-fa21-003.netlify.app/appex/data/yawn-sim.csv")
```

```{r sim-dist, eval = F}
ggplot(data = sim_data, mapping = aes(x = diff_in_prop)) +
  geom_histogram(binwidth = 0.05) + 
  labs(title = "Your Results: Difference in Proportion of Yawners")
```

- What is the approximate center of the distribution? Is this what you expected? Why or why not?

- The observed difference in proportions from the Mythbusters episode is 0.0441. Based on your simulated distribution, do yawning and seeing someone yawn appear to be dependent?

## Exercise 3 

Let's use the data from the *Mythbusters* episode and simulation-based inference in R to test this claim. Based on their experiment, do yawning and seeing someone yawn appear to be dependent?

Evaluate this question using a simulation based approach. We will use the same null and alternative hypotheses as before. The data from *Mythbusters* is available in the `yawn` data frame. 

- Fill in the code below to generate the null distribution. Uncomment the code once it is complete.

```{r mythbusters-null}
set.seed(101821)
#null_dist <- yawn %>%
#  specify(response = ____, explanatory = _____, success = "yawn") %>%
#  hypothesize(null = "______") %>%
#  generate(100, type = "permute") %>%
#  calculate(stat = "______", 
#            order = c("trmt", "ctrl"))
```

- Visualize the null distribution and shade in the area used to calculate the p-value. 

```{r mythbusters-null-viz}
# add code 
```

- Calculate p-value. Then use the p-value to make your conclusion using a significance level of 0.1.

```{r mythbusters-calc-p-value}
# add code

```

## Exercise 5

Do you buy the conclusions from this experiment? Why or why not? 


<br>

<hr> 

[^1]: This question was adapted from [*Introduction to Modern Statistics*](https://openintro-ims.netlify.app/).
[^2]: Simulation activity from [Data science in a box](https://datasciencebox.org/)